---
title: "Exploratory Analysis of Crime Offences in European Countries (2021)"
author: "Divya Pariti"
format: pdf
editor: visual
---

### Assignment 1

The data set `EurostatCrime2021.xlsx` contains the records of offences by offence category in 41 European Countries in 2021.

## Task - 1 : Manipulation

Q1. Load the data set `EurostatCrime2021.xlsx`:

```{r}
# Loading the `readxl` package
library(readxl)
# Loading the data from the Excel sheet
data = read_excel("EurostatCrime2021.xlsx", skip = 6)
#Viewing the first few rows by using `head()` function
head(data)
```

Q2. What is the size (number of rows and columns) and the structure of this data set?

```{r}
# Finding the size of the data set
size = dim(data)
# `dim()` function sets the dimension of the given data frame
# Printing the number of rows
print(paste("The number of rows in the dataset:", size[1]))
# Printing the number of columns
print(paste("The number of colummns in the dataset:", size[2]))

# Displaying the structure of the data set
str(data)
#`str()` displays the structures of objects or the summary of the output produced.

```

Q3. Remove the columns Fraud and Money laundering (they contain no data).

```{r}
# Using `subset()` function to perform the above operation on the data
new_data = subset(data, select = -c(`Fraud`, `Money laundering`))
# We can use `subset()` function to select parts of the matrix, vector or data frame.
print(new_data)
```

Q4. For some countries Theft includes also Theft of a motorized vehicle or parts thereof, Burglary, and Burglary of private residential premises, in others they are recorded separately. To compare different countries, remove the columns involving theft and burglary: • Theft, • Theft of a motorized vehicle or parts thereof, • Burglary, • Burglary of private residential premises

```{r}
new_data1 = subset(new_data, select = -c(
  Theft,
  `Theft of a motorized vehicle or parts thereof`,
  Burglary,
  `Burglary of private residential premises`
))
print(new_data1)
head(new_data1)

```

Q5. Add a column containing the overall record of offences for each country (per hundred thousand inhabitants)

```{r}
# Using the set `new_data1` to add columns to the overall record of offences
# Ensuring that all the columns contain numeric data and so converting the columns to numeric using `as.numeric()` function
new_data1[, -1] = lapply(new_data1[, -1], as.numeric)
new_data1$Record = rowSums(new_data1[, -1], na.rm=TRUE)
new_data1$Record

```

Q6. Work with the data set you just created, and write some code to list the countries that contain any missing data.

```{r}
new_data1$Country[apply(is.na(new_data1), 1, any)]

```

Q7. Remove the countries with missing data.

```{r}
# Using `na.omit()` function 
# This function removes rows with missing values from the data set
updated_data = na.omit(new_data1)
updated_data

```

Q8. How many observations and variables are in this new data set?

```{r}
# Using the `updated_data` to find out observations and variables
updated_size = dim(updated_data)

# Printing number of rows
print(paste("The number of rows in the new data set are:", updated_size[1]))

# Printing number of columns
print(paste("The number of columns in the new data set are:", updated_size[2]))
```

## Task - 2: Analysis

Working with the data set produced at the end of Task 1

Q1. Which country has the highest overall record of offences in 2021 (per hundred thousand inhabitants)? To get full marks you must also provide the R code that returns that country name only.

```{r}
new_data3 = updated_data[order(updated_data$Record, decreasing = TRUE), ]

# Printing the data of country with highest record of offences in 2021 (per hundred thousand inhabitant)
print(paste("The country with highest record of offences in 2021 is : ",new_data3$Country[1]))
```

Q2. Produce a table showing the countries and the proportion of the overall crimes due to acts against computer system, sort the rows in ascending order of proportions, and display only the first three decimal digits.

```{r}
# Creating the table to produce list of countries and their proportion of the overall crimes due to acts against computer system.
updated_data$`Acts against computer systems` = round(updated_data$`Acts against computer systems` / updated_data$Record, 3)

# Creating the table using data frames
crimes_table = data.frame(Country = updated_data$Country, Proportion = updated_data$`Acts against computer systems`)

# sorting the rows and displaying the first three decimal digits.
sorted_crimes_table = crimes_table[order(crimes_table$Proportion), ]

# Printing the sorted table with three decimal digits
sorted_crimes_table

```

Q3. Create a plot displaying the relationship between robbery and unlawful acts involving controlled drugs or precursors. Make the plot “nice” i.e., show country names, change size of the plot, change axis labels, etc.

```{r}

# Loading the package "ggplot2"
library(ggplot2)

# Plotting the graph to display the relationship
ggplot(updated_data, aes(x = Robbery, 
                         y = `Unlawful acts involving controlled drugs or precursors`, label = Country)) +
  geom_point(size = 2) +
  geom_text(hjust = 0.5, vjust = -0.6) +
  labs(x = "Robbery", y = "Controlled Drugs or Precursors") +
  ggtitle("Country wise Robbery vs Controlled Drugs/Precursors")

```

## Task 3 Creativity

Do something interesting with these data! Create two plots showing something we have not discovered above already and outline your findings. For this Task you can decide if you want to use the original data set from Task 1 Question 1 or the modified one.

plot 1

```{r}
# Loading the required package `ggplot2`
library(ggplot2)

# Comparing the `Sexual Exploitation` data of all the countries present in the given data set and plotting a bar graph
ggplot(updated_data, aes(x = reorder(Country, `Sexual exploitation`), y = `Sexual exploitation`, fill = Country)) +
  geom_bar(stat = "identity") +
  labs(x = "Country", y = "Sexual Exploitation Data") +
  theme_minimal() +
  ggtitle("Sexual Exploitation Data of all countries")

updated_data

```

For the above bar graph, we have used the cleaned and updated version of the data set. We have compared the `Sexual Exploitation` between countries from the data set and we can conclude that `Austria` has the highest amount of `Sexual Exploitation` and `Greece` has the lowest.

plot 2

```{r}

# For the second plot let us look at the highest assault rates and highlight the countries
# Loading the required package
library(ggplot2)

# Sorting the data by Sexual Assault rate in descending order
sorted_data = updated_data[order(updated_data$`Sexual assault`, decreasing = TRUE), ]

# Creating a bar plot for countries with the highest Sexual Assault rates
countries = head(sorted_data, 15)
ggplot(countries, aes(x = reorder(Country, -`Sexual assault`), y = `Sexual assault`)) +
  geom_bar(stat = "identity") +
  labs(x = "Country", y = "Sexual assault") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Countries with the Highest Assault Rates")

updated_data

```

For the above bar graph, we have used the cleaned and updated version of the data set. We have compared the `Sexual Assault` between countries from the data set and we can conclude that `Finland` has the highest amount of `Sexual Assault` and again `Greece` has the lowest rates in both `Sexual Assault` as well as `Sexual Exploitation`.
